---
title: "Comparing Model Implementations (1) - Tables"
author: "Khaled Fouda"
date: "`r Sys.Date()`"
output:
  html_document:
     df_print: paged 
     output_file: Documentation/html/Comapring_Model_Implementations_Tables.html
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=TRUE, echo=F, message=FALSE, warning=FALSE}
library(kableExtra)
library(magrittr)
library(tidyverse)
root_dir = "/mnt/campus/math/research/kfouda/main/HEC/Youssef/HEC_MAO_COOP/"
knitr::opts_knit$set(root.dir=root_dir)
knitr::opts_chunk$set(echo = FALSE, cache = FALSE, eval = TRUE, 
                      cache.path = paste0(root_dir,"/Documentation/cache/"))
#source("./Mao/SMC_functions6.R")
```


```{r include=TRUE, message=FALSE, warning=FALSE}
source("./code_files/import_lib.R")
#path_to_data = "./Mao/saved_data/"
library(data.table)
rmarkdown::render("Documentation/Rmd//4_Model_Comparison(1)_Tables.Rmd",output_dir = "Documentation/html/")
```

### Model Summary

```{r}
Models <- c(
   "Mao",
   "SoftImpute",
   "CAMC_holdout",
   "CAMC_kfold",
   "CASMC_holdout",
   "CASMC_kfold",
   "Naive"
)

Note <- c(
   "Mao's original model",
   "Soft-Impute (fit function from library and CV from Vignette)",
   "Covariate-Adjusted Matrix Completition with holdout cross-validation",
   "Covariate-Adjusted Matrix Completition with K-fold cross-validation",
   "Covariate-Adjusted Sparse  Matrix Completition (Sparse-plus-low-rank) with holdout cross-validation",
   "Covariate-Adjusted Sparse  Matrix Completition (Sparse-plus-low-rank) with K-fold cross-validation",
   "Missing Values are replaced with the middle point of row and column average."
)

model_rep <-
   c(
      "$\\begin{aligned} O = X \\beta + M  \\end{aligned}$",
      "$\\begin{aligned} O =  M  \\end{aligned}$",
      "$\\begin{aligned} O = X \\beta + AB^T  \\end{aligned}$",
      "$\\begin{aligned} O = X \\beta + AB^T  \\end{aligned}$",
      "$\\begin{aligned} O = P_{\\Omega}(O-X \\beta - AB^T) + (X \\beta + AB^T)\\\\ =Sparse +  Low Rank  \\end{aligned}$",
      "$\\begin{aligned} O = P_{\\Omega}(O-X \\beta - AB^T) + (X \\beta + AB^T)\\\\ =Sparse +  Low Rank\\end{aligned}$",
      ""
   )

hyperparameters <- c(
   "$(\\Theta,\\lambda_1, \\lambda_2,\\alpha)$",
   "$(\\lambda_2,J)$",
   "$(\\lambda_1, \\lambda_2, J)$",
   "$(\\lambda_1, \\lambda_2, J)$",
   "$(\\lambda_2, J)$",
   "$(\\lambda_2, J)$",
   ""
)

include_cov <- rep("Yes", 7)
include_cov[c(2, 7)] <- "No"
Cross_validation_type <- rep("80%/20%", 6, 7)
Cross_validation_type[c(1)] <- "K-fold (5)"
Cross_validation_type[c(4)] <- "K-fold (3)"
Cross_validation_type[c(6)] <- "K-fold (5)"
Cross_validation_type[c(7)] <- ""
Regular_cov <- rep("Ridge", 7)
Regular_cov[2] <- "None"
Regular_cov[7] <- ""
Regular_cov[5:6] <- "Dimension Reduction"
New_Model <- rep("Yes", 7)
New_Model[1:2] <- "No"
New_Model[7] <- ""

data.frame(
   Model = Models,
   Note = Note,
   `Has Covariates?` = include_cov,
   `Cross-Validation Type` = Cross_validation_type,
   `Regularization on the Covariates?` = Regular_cov,
   `New Model` = New_Model,
   Model.Representation = model_rep,
   Hyperparameters = hyperparameters
) %>%
   kable("html",
         escape = FALSE,
         #align = 'c',
         caption = "Model Summary Table; $Y= P_{\\Omega}(O) + \\epsilon$ is observed where $\\Omega$ are the observed indices and $\\epsilon$ is a small random noise. (ie, Y is partially observed and slightly corrupted version of O.") %>%
   kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = F,
      position = "center"
   ) %>% 
   column_spec(2, width = "300px")

```



```{r}
file_list <- list.files(path = path_to_data, pattern =
                           "((Compare_MC_Models_Youssef_Simulation_[0-9]+.*)|(Compare_MC_Models_Mao_Simulation))\\.csv$",
                        full.names = TRUE,recursive =F)

results = data.frame()
for(f in file_list)
   results = rbind(results, read.csv(f))

results %<>%
   mutate(simulation = case_when(missinginess == 0 ~ "Mao",
                                 missinginess == 0.8 & collinearity == TRUE ~ "80% Missing with Coll",
                                 missinginess == 0.8 & collinearity == FALSE ~ "80% Missing",
                                 missinginess == 0.9 & collinearity == TRUE ~ "90% Missing with Coll",
                                 missinginess == 0.9 & collinearity == FALSE ~ "90% Missing"))

```



```{r}
# Function to color minimum values in green



color_min_values <- function(data, columns) {
   data <- as.data.frame(data)
   for (col_name in columns) {
      col_min <- min(as.numeric(data[[col_name]]), na.rm = TRUE)
      colrounded <- round(as.numeric(data[[col_name]]), 4)
      data[[col_name]] <-
         cell_spec(
            ifelse(as.numeric((is.na(
               data[[col_name]]
            ))), '', colrounded),
            "html",
            color = ifelse(as.numeric((is.na(
               data[[col_name]]
            ))), "grey98", "black"),
            background = fifelse((data[[col_name]]) == col_min, "lightgreen", "grey98", "grey98")
         )
   }
   col_min <- min(abs(data[["rank_diff"]]), na.rm = TRUE)
   data[["rank"]] <- cell_spec(data[["rank"]],
                               "html",
                               background = ifelse(abs(data[["rank_diff"]]) == col_min, "lightgreen", "grey98"))
   return(data)
}
# results <- results_MaoSim
print_table <-
   function(results,
            missingness = NULL,
            coll = TRUE,
            prob = NA,
            caption = "",
            nmodels = 7) {
      #$$
      names(results) = gsub("error\\.", "", names(results))
      results %>%
         filter(missinginess == missingness, collinearity == coll) %>%
         dplyr::rename(
            O = all,
            O_test = test,
            Method = model,
            missing_prob = missinginess,
            seconds = time,
            M = B,
            Beta = beta
         ) %>%
         mutate(rank_diff = rank - true_rank) %>%
         select(
            Method,
            true_rank,
            missing_prob,
            lambda.1,
            lambda.2,
            O_test,
            O,
            M,
            Beta,
            rank,
            rank_diff,
            seconds
         )   -> combined_results
      combined_results %<>% 
         mutate(seconds = ifelse(Method == "Naive", NA, seconds))
      
      #$$
      test_col = 6
      if (missingness == 0) {
         test_col = 5
         combined_results %<>% select(-missing_prob)
         nfixed = 2
      } else
         nfixed = 3
      
      
      # Color the minimum errors for each pair of rows (200x200, 400x400, etc.)
      error_columns <- c("O_test", "O", "Beta", "M", "seconds")
      for (i in seq(1, 28, nmodels))
         combined_results[i:(i + nmodels - 1),] <-
         color_min_values(combined_results[i:(i + nmodels - 1),], error_columns)
      
      combined_results %>%
         mutate_if(is.numeric, function(x)
            replace_na(as.character(round(x, 4)), '')) %>%
         
         kable("html",
               escape = FALSE,
               align = 'c',
               caption = caption) %>%
         kable_styling(
            bootstrap_options = c("striped", "hover", "condensed"),
            full_width = F,
            position = "center"
         ) %>%
         add_header_above(c(
            "Fixed" = nfixed,
            "Hyperparameters" = 2,
            "Error on" = 4,
            "Rank" = 2,
            "Time" = 1
         )) %>%
         pack_rows("400x400", 1, nmodels) %>%
         pack_rows("600x600", (nmodels * 1) + 1, nmodels * 2) %>%
         pack_rows("800x800", (nmodels * 2) + 1, nmodels * 3) %>%
         pack_rows("1000x1000", (nmodels * 3) + 1, nmodels * 4) %>%
         column_spec(
            test_col,
            bold = T,
            background = grDevices::adjustcolor("red", alpha.f = 0.1)
         )
      
      
   }
```


## Simulation 1:   Mao's Simulation; 10 Covariates; 90% Missing; No Collinearity 

```{r fig.height=15, fig.width=15}
print_table(results,0,F, caption = "Data was simulated using Mao with m=10 and r=10")
```

## Simulation 2:   Yousef's Simulation; 10 Covariates; 80% Missing; No Collinearity

```{r fig.height=15, fig.width=15}
print_table(results,0.8,F, "Data was simulated using Youssef's model with m1=m2=10, 80% missing values, and without collinearity")
```

## Simulation 3:   Yousef's Simulation; 10 Covariates; 80% Missing; With Collinearity

```{r fig.height=15, fig.width=15}
print_table(results,0.8,T, "Data was simulated using Youssef's model with m1=m2=10, 80% missing values, and with collinearity")
```

## Simulation 4:   Yousef's Simulation; 10 Covariates; 90% Missing; No Collinearity

```{r fig.height=15, fig.width=15}
print_table(results,0.9,F, "Data was simulated using Youssef's model with m1=m2=10, 90% missing values, and without collinearity")
```

## Simulation 5:   Yousef's Simulation; 10 Covariates; 90% Missing; With Collinearity

```{r fig.height=15, fig.width=15}
print_table(results,0.9,T, "Data was simulated using Youssef's model with m1=m2=10, 90% missing values, and with collinearity")
```